---
output:
  html_document:
    df_print: paged
    theme: cerulean
    highlight: haddock
    toc: yes
    toc_float: yes
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE, 
                      fig.width = 10, 
                      fig.height = 6)

```

<br>

# Data

View the dataset documentation here: [Syracuse, NY Tax Parcel Data](https://watts-college.github.io/cpp-526-fall-2022/labs/syr_parcels.html).

<br>

```{r}

library(sp)
library(rgdal)
library(geojsonio)                            # Load required libraries

```

<br>

## Import Spatial Polygons

<br>

```{r cache = TRUE}

url <- paste0("https://raw.githubusercontent.com/DS4PS/Data",
              "-Science-Class/master/DATA/downtown-syr.geojson")

downtown <- geojson_read(url, what = "sp")    # Load parcel data

plot(downtown,  
     border = "gray60",
     col = "gray80")                          # Visualize parcels

```

<br>
<br>

# Example Solutions

### Example: Tax Parcel Acres

Parcels with more than one acre

<br>

```{r}

result <- downtown$acres > 1                  # Each value in 'acres' is tested

```

<br>

```{r}

group.colors <- ifelse(test = result,         # 'test =' accepts TRUE or FALSE values
                       yes = "firebrick",     # 'yes =' converts values that are TRUE
                       no = "gray80")         # 'no =' converts values that are FALSE

```

<br>

```{r}

plot(downtown,                                # Plots object 'downtown'
     border = NA,
     col = group.colors)                      # Red when TRUE; gray when FALSE

```

<br>
<br>

### Example: Parcels with Single Families
<br>

```{r}

unique(downtown$landuse)                      # All unique values in 'landuse'

```

<br>

```{r}

result <- downtown$landuse == "Single Family" # Test if 'landuse' is "Single Family"

group.colors <- ifelse(test = result,         # Provide TRUE or FALSE test results
                       yes = "firebrick",     # If TRUE, make "firebrick"
                       no = "gray80")         # If FALSE, make "gray80"

plot(downtown,
     border = NA, 
     col = group.colors)                      # Plot with respective colors

```

<br>

# Questions & Solutions

## Question 1: Vacant Lots

<br>

```{r}

unique(downtown$landuse)                      # Print unique values in 'landuse'

```

<br>

```{r}

result <- downtown$landuse == "Vacant Land"

group.colors <- ifelse(test = result, 
                       yes = "firebrick", 
                       no = "gray80")

plot(downtown,
     border = NA, 
     col = group.colors)

```

<br>
<br>

## Question 2: Parking Lots

<br>

```{r}

# Use function: 'sum()'
# Use variable: 'landuse'

# Assign test (logical) output to object 'result' by replacing 'FALSE'


result <- downtown$landuse == "Parking" 
parking_lots <- sum(result, na.rm=T)
cat("The number of parking lots is ", parking_lots)
group.colors <- ifelse(test = result, 
                       yes = "firebrick", 
                       no = "gray80")

plot(downtown,
     border = NA, 
     col = group.colors)

```

<br>
<br>

## Question 3: New Construction

<br>

```{r}

# Use variable: 'yearbuilt'
# Find the number of buildings constructed either after or during and after than 1980

# Assign test (logical) output to object 'result' by replacing 'FALSE'
result <- downtown$yearbuilt >= 1980 
num_buildings <- sum(result, na.rm=T)
cat("The number of buildings constructed after 1980 is ", num_buildings)

group.colors <- ifelse(test = result, 
                       yes = "firebrick", 
                       no = "gray80")

plot(downtown,
     border = NA, 
     col = group.colors)

```

<br>

```{r}

# Use variable: 'landuse'
# Use variable: 'yearbuilt'
# Count only commercial properties with buildings constructed during or later than 1980
commercial_after1980 <- downtown$landuse=="Commercial" & downtown$yearbuilt >= 1980
x <- sum(commercial_after1980, na.rm=T ) / length(commercial_after1980)

cat("The proportion of commercial properties built since 1980 is ", x , "%.\n")

```

<br>
<br>

## Question 4: Parcels Valued Over $10M

<br>

```{r}

# Use function: 'sum()'
# Use variable: 'assessedval'

# Assign test (logical) output to object 'result' by replacing 'FALSE'
result <- downtown$assessedval > 10000000 
val_10m <- sum(result, na.rm=T)
cat(val_10m, "parcels have assessed values over $10 M.")
group.colors <- ifelse(test = result, 
                       yes = "firebrick", 
                       no = "gray80")

plot(downtown,
     border = NA, 
     col = group.colors)

```

<br>

## Question 5: Total Tax Delinquent Parcels

<br>

```{r}

# Use function: 'sum()'
# Use variable: 'amtdelinqt'

# Assign test (logical) output to object 'result' by replacing 'FALSE'

result <- downtown$amtdelinqt > 0
delinquent <- sum(result, na.rm=T)
cat(delinquent, "properties are delinquent on taxes.")

group.colors <- ifelse(test = result, 
                       yes = "firebrick", 
                       no = "gray80")

plot(downtown,
     border = NA, 
     col = group.colors)

```

<br>
<br>

## Question 6: Tax Delinquent Parcels
<br>

```{r}

# Use function: 'mean()'
# Use variable: 'amtdelinqt'

result <- downtown$amtdelinqt > 0
delinquent_p <- mean(result, na.rm=T)
cat(delinquent_p, "% of parcels have delinquent tax payments owed.")

```
<br>

## Question 7: Tax Delinquent Commercial Properties

**Answer I:** **[X]%** of commercial properties are delinquent on taxes. 

**Answer II:** **[X]%** of delinquent tax dollars are owed by commercial parcels.

<br>

```{r}

# Use function: 'mean()'
# Use variable: 'amtdelinqt'
# Use variable: 'landuse'

commercial_properties <- downtown$landuse == "Commercial" &  downtown$amtdelinqt > 0
proportion_delinquent <- mean(commercial_properties, na.rm = TRUE)

cat(proportion_delinquent, "% of commercial properties are delinquent on taxes.")

# The first answer is tax-delinquent commercial properties over all commercial properties
# The second answer is the tax dollars owed by commercial properties (a subset) over all tax dollars owed

```
<br>

## Question 8: Tax Delinquent Parcels by Land Use
<br>

```{r}

# Use function: 'table()'
# Use variable: 'amtdelinqt'
# Use variable: 'landuse'

delinquent_properties <- downtown[downtown$amtdelinqt > 0, ]
delinquent_by_landuse <- table(delinquent_properties$landuse)
print(delinquent_by_landuse)
```


